---
title: "📊 Statistical Analysis  "
output:
  html_document: 
    code_folding: show
    toc: yes
    toc_depth: 5
    toc_float: yes
    fig_width: 9
    fig_height: 6
---

```{r setup, include=FALSE}
# To show or hide the code in html
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, include=FALSE}
# Data manip
library(dplyr)
library(broom)
library(tidyr)
library(readr)
library(tibble)
library(mgsub)
library(tidyverse)
library(DT)
library(sjPlot)

# Stats
library(glmmTMB)
library(lmerTest)
library(lme4)
library(MuMIn)
library(rstatix)

# Data viz
library(gridExtra)
library(kableExtra)
library(knitr)
library(ggplot2)
library(ggpubr)
```

# 🎯 Objective  
This RMarkdown file is designed to perform statistical analyses of the dataset.  
It includes steps such as importing the dataset, selecting conditions, analyzing variable distributions, and examining statistical relationships between tha variables.  

The analysis can focus on one of two conditions:  
- **Odor Recognition**  
- **Associative Memory**  

📌 **Important:** The condition selection is done in the step **[Condition Selection](#condition-selection)**, where the user must specify the appropriate condition before proceeding with the analysis.  

# 📑 Table of Contents  

- **[Import the Dataset](#import-the-dataset)**  
  - *Load the dataset for analysis.*  
  - [Dataset by Trial](#dataset-by-trial) – *Explore the dataset at the trial level.*  
  - [Dataset by Subject](#dataset-by-subject) – *Aggregate data at the participant level.*  
  - [Condition Selection](#condition-selection) – *Choose the condition of interest for the analysis.*  
  - [Variable Distribution](#variable-distribution) – *Visualize the distribution of key variables.*  
- **[Perceptual Scores](#perceptual-scores)**  
  *Analyze perceptual-related measures across conditions.*  
- **[Gender × Memory Scores](#gender-memory-scores)**  
  *Examine how memory scores vary across genders.*  
- **[Gender × Features](#gender-features)**  
  *Explore relationships between gender and all the features.*  
- **[Memory Score × Features](#memory-score-features)**  
  *Assess the statistical effect of each independent feature on memory scores.* 

# Import the dataset

```{r loading}
# Get the parent directory of the current working directory
project_dir <- dirname(dirname(getwd()))

# Define the path to the data folder
data_folder <- file.path(project_dir, "FINAL/data")

# Define the path to the CSV file
file_path <- file.path(data_folder, "dataset.csv")

# Read the CSV file
df <- read.csv(file_path)
```

```{r preprocessing}
# Separate pleasantness into unpleasant and pleasant
df$unpleasant <- ifelse(
  df$pleasantness >= -5 & df$pleasantness <= 0,
  df$pleasantness,
  NA
                        )

df$pleasant <- ifelse(
  df$pleasantness >= 0 & df$pleasantness <= 5,
  df$pleasantness,
  NA
                      )

# Reorder columns
col_order <- c(
  'study', 'participant', 'gender', 'gender_encoded',
  'pres_order', 'odor_num', 'odor_name', 'is_target', 'day',
  'hit', 'cr', 'www', 'wwhich','mem',
  'pleasantness', 'unpleasant', 'pleasant',
  'emotional_strength', 'intensity', 'familiarity',
  'avg_distance_target', 'avg_distance_hit',
  "descriptors", "tokens", "lemma",
  "nb_words",
  "mean_lemma_jaccard_target", "mean_lemma_jaccard_hit"
  )

# Select col
df_all <- df %>% select(all_of(col_order))

# Subset df in target odors
df_target <- filter(df, is_target == 1)

# Subset df in hit odors
df_hit <- filter(df_target, hit == 1)
```

## Dataset by trial

```{r display df}
# Display df
datatable(
  df_all,
  options = list(
    scrollX = TRUE,
    scrollY = "400px"
    )
  )
```

```{r stats summary}
# Define the columns to include
selected_cols_all <- c(
  'pleasantness',
  'emotional_strength',
  'intensity',
  'familiarity',
  "nb_words"
)

# Function to compute mean and standard deviation for selected columns
summary_stats <- function(df, cols) {
  df %>%
    select(all_of(cols)) %>%
    summarise(
      across(everything(),
             list(
               mean = ~mean(.x, na.rm = TRUE),
               sd = ~sd(.x, na.rm = TRUE)
               )
             )
      ) %>%
    pivot_longer(
      everything(),
      names_to = c("variable", ".value"),
      names_pattern = "(.+)_(mean|sd)$"
      )
}

# Compute statistics for each dataset
stats_all <- summary_stats(
  df_all,
  selected_cols_all
  ) %>%
  mutate(group = "df_all")

stats_target <- summary_stats(
  df_target, 
  c(selected_cols_all,
    intersect(c(
      'hit',
      'avg_distance_target',
      'mean_lemma_jaccard_target'
      ),
      names(df)
              )
    )
) %>%
  mutate(group = "df_target")

stats_hit <- summary_stats(
  df_hit, 
  c(selected_cols_all,
    intersect(c(
      'www',
      'wwhich',
      'mem',
      'avg_distance_hit',
      'mean_lemma_jaccard_hit'),
      names(df)
              )
    )
) %>%
  mutate(group = "df_hit")

# Combine results into a single table
final_stats <- bind_rows(stats_all, stats_target, stats_hit) %>%
  mutate(mean_sd = paste0(round(mean, 2), " ± ", round(sd, 2))) %>%
  select(group, variable, mean_sd) %>%
  pivot_wider(names_from = variable, values_from = mean_sd)

# Display df
datatable(
  final_stats,
  options = list(
    scrollX = TRUE,
    scrollY = "400px"
    )
  )
```

## Dataset by subj

```{r scores by subj}
# Compute scores by subj
by_subj <- df_all %>%
  group_by(participant) %>%
  summarise(
    gender = first(gender),
    study = first(study),
    Hit = sum(hit),
    cr = sum(cr),
    www = sum(www),
    wwhich = sum(wwhich),
    mem = sum(mem),
    n_target = sum(is_target == 1),
    n_distractor = sum(is_target == 0),
    pleasantness_t = mean(pleasantness[is_target == 1], na.rm = TRUE),
    pleasantness_h = mean(pleasantness[hit == 1], na.rm = TRUE),
    emotional_strength_t = mean(emotional_strength[is_target == 1], na.rm = TRUE),
    emotional_strength_h = mean(emotional_strength[hit == 1], na.rm = TRUE),
    intensity_t = mean(intensity[is_target == 1], na.rm = TRUE),
    intensity_h = mean(intensity[hit == 1], na.rm = TRUE),
    familiarity_t = mean(familiarity[is_target == 1], na.rm = TRUE),
    familiarity_h = mean(familiarity[hit == 1], na.rm = TRUE),
    nb_words_t = mean(nb_words[is_target == 1], na.rm = TRUE),
    nb_words_h = mean(nb_words[hit == 1], na.rm = TRUE)
  ) %>%
  mutate(
    hit_prop = (Hit) / (n_target),
    www_prop = (www) / (n_target),
    score_www = (www) / (Hit),
    score_mem = (mem) / (Hit),
    HR = (Hit + 0.5) / (n_target + 1),
    FR = ((n_distractor - cr) + 0.5) / (n_distractor + 1),
    d_prime = log((HR * (1 - FR)) / (FR * (1 - HR)))
  )

# Display df
datatable(
  by_subj,
  options = list(
    scrollX = TRUE,
    scrollY = "400px"
    )
  )
```

```{r stats summary by_subj}
# Define the columns to include
selected_cols_all <- c(
  'hit_prop',
  'd_prime',
  'score_www',
  'score_mem',
  'pleasantness_t',
  'pleasantness_h',
  'emotional_strength_t',
  'emotional_strength_h',
  'intensity_t',
  'intensity_h',
  'familiarity_t',
  'familiarity_h',
  "nb_words_t",
  "nb_words_h"
)

# Compute statistics
stats_by_subj <- summary_stats(by_subj, selected_cols_all) %>% mutate(group = "by_subj")

# Display df
datatable(
  stats_by_subj,
  options = list(
    scrollX = TRUE,
    scrollY = "400px"
    )
  )
```

## Condition selection

```{r condition selection}
# Select a condition
condition <- "odor_recognition" # "odor_recognition" or "associative_memory"

# Initialize variables according to condition
if (condition == "odor_recognition") {
  data <- df_target
  data$outcome <- data$hit
  by_subj$outcome <- by_subj$hit_prop
  data$percept_dist <- data$avg_distance_target
  data$jaccard_dist <- data$mean_lemma_jaccard_target
  fillcolor <- "#5187EC"
  chance <- 1/2
}

if (condition == "associative_memory") {
  data <- df_hit
  data$outcome <- data$mem
  by_subj$outcome <- by_subj$score_mem
  data$percept_dist <- data$avg_distance_hit
  data$jaccard_dist <- data$mean_lemma_jaccard_hit
  fillcolor <- "#5EACA3"
  chance <- 1/3
}
```

# Variable distribution {.tabset .tabset-fade}

```{r custom theme for plots}
linewidth = 1
textsize = 25

custom_theme <- function() {
  theme_bw() +
    theme(
      panel.background = element_rect(fill = "white"),
      panel.border = element_blank(),
      aspect.ratio = 1/1,
      axis.line = element_line(
        color = "black",
        linewidth = linewidth
        ),
      axis.text.x = element_text(
        color = "black",
        size = textsize
        ),
      axis.text.y = element_text(
        color = "black", size = textsize),
      axis.title = element_text(
        color = "black",
        size = textsize
        ),
      axis.ticks = element_line(
        color = "black",
        linewidth = linewidth
        ),
      axis.ticks.length = unit(10, "pt"),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      legend.position = "none",
      plot.background = element_rect(fill = "white")
    )
}
```

## Demographics

```{r demographics distribution, fig.height=5, fig.width=15}
# Study distribution
study <- ggplot(
  by_subj,
  aes(x = factor(study))
  ) +
  geom_bar(
    aes(
      y = after_stat(count)),
      fill = c("#d8bfbf", "#d4a6a6", "#a074b7", "#4d2654")
    ) +
  geom_text(
    stat = 'count',
    aes(label = after_stat(count)),
    vjust = -0.2
    ) +
  labs(
    x = "Study",
    y = "Number of Participants"
    ) +
  custom_theme()

# Gender distribution
gender <- ggplot(by_subj, aes(x = gender)) +
  geom_bar(
    aes(y = after_stat(count)),
    fill = c("#C08FA0", "#8FC0AE")
    ) +
  geom_text(
    stat = 'count',
    aes(label = after_stat(count)),
    vjust = -0.2
    ) +
  labs(
    x = "Gender",
    y = "Number of Participants"
    ) +
  custom_theme()

# Display plots
ggarrange(study, gender,
          nrow = 1,
          ncol = 3
)
```

## Memory variable

```{r memory distribution, fig.height=5, fig.width=10}
if (condition == "odor_recognition") {
  x_label_prop <- "Hit proportion"
  x_label_dist <- "Odor Recognition"
  labels_outcome <- c("0" = "Miss", "1" = "Hit")
  limits <- c(0,800)
  breaks <- seq(0,800,by=200)
} else if (condition == "associative_memory") {
  x_label_prop <- "Correct retrieval proportion"
  x_label_dist <- "Associative Memory"
  labels_outcome <- c("0" = "Incorrect", "1" = "Correct")
  limits <- c(0,600)
  breaks <- seq(0,600,by=200)
}

# Outcome distribution (proportion)
outcome_prop <- ggplot(by_subj, aes(x = outcome)) +
  geom_histogram(
    binwidth = 0.1,
    fill=fillcolor,
    color="black",
    aes(y = after_stat(count))
    ) +
  scale_x_continuous(
    limits=c(-0.05,1.05),
    breaks=seq(0,1,by=0.2)
    ) +
  scale_y_continuous(
    limits=c(0,35),
    breaks=seq(0,35,by=5)
    ) +
  labs(
    x = x_label_prop,
    y = "Number of Participants"
    ) +
  custom_theme()

# Outcome distribution (bar plot)
outcome_dist <- ggplot(data, aes(x = factor(outcome))) +
  geom_bar(
    fill=fillcolor,
    color="black",
    aes(y = after_stat(count))
    ) +
  scale_x_discrete(
    labels = labels_outcome
    ) +
  scale_y_continuous(
    limits=limits,
    breaks=breaks
    ) +
  labs(
    x = x_label_dist,
    y = "Number of Trials"
    ) +
  custom_theme()

# Display plots
ggarrange(outcome_prop, outcome_dist,
          ncol = 2
)
```

## Perceptual variables

```{r percptual variables distribution, fig.height=5, fig.width=20}
# Pleasantness distribution
pleasantness <- ggplot(data, aes(x = pleasantness)) +
  geom_histogram(
    binwidth = 1,
    fill=fillcolor,
    color="black",
    aes(y = after_stat(count))
    ) +
  scale_x_continuous(
    limits=c(-5.5,5.5),
    breaks=seq(-5,5, by=2.5)
    ) +
  scale_y_continuous(
    limits=c(0,200),
    breaks=seq(0,200,by=50)
    ) +
  labs(
    x = "Pleasantness",
    y = "Number of Trials"
    ) +
  custom_theme()

# emotional_strength distribution
emotional_strength <- ggplot(data, aes(x = emotional_strength)) +
  geom_histogram(
    binwidth = 1,
    fill=fillcolor,
    color="black",
    aes(y = after_stat(count))
    ) +
  scale_x_continuous(
    limits=c(-0.5,5.5),
    breaks=seq(0,5, by=1)
    ) +
  scale_y_continuous(
    limits=c(0,300),
    breaks=seq(0,300,by=100)
    ) +
  labs(
    x = "Emotional strength",
    y = "Number of Trials"
    ) +
  custom_theme()

# Intensity episodic distribution
intensity <- ggplot(data, aes(x = intensity)) +
  geom_histogram(
    binwidth = 1,
    fill=fillcolor,
    color="black",
    aes(y = after_stat(count))
    ) +
  scale_x_continuous(
    limits=c(-0.5,10.5),
    breaks=seq(0,10, by=2.5)
    ) +
  scale_y_continuous(
    limits=c(0,200),
    breaks=seq(0,200,by=50)
    ) +
  labs(
    x = "Intensity",
    y = "Number of Trials"
    ) +
  custom_theme()

# Familiarity distribution
familiarity <- ggplot(data, aes(x = familiarity)) +
  geom_histogram(
    binwidth = 1,
    fill=fillcolor,
    color="black",
    aes(y = after_stat(count))
    ) +
  scale_x_continuous(
    limits=c(-0.5,10.5),
    breaks=seq(0,10, by=2.5)
    ) +
  scale_y_continuous(
    limits=c(0,200),
    breaks=seq(0,200,by=50)
    ) +
  labs(
    x = "Familiarity",
    y = "Number of Trials"
    ) +
  custom_theme()

# Display plots
ggarrange(pleasantness, emotional_strength, intensity, familiarity,
          nrow = 1,
          ncol = 4
)
```

## Perceptive distance

```{r perceptive distance distribution, fig.height=5, fig.width=5}
# Perceptive distance distribution
dist <- ggplot(data, aes(x = percept_dist)) +
  geom_histogram(
    binwidth = 1,
    fill=fillcolor,
    color="black",
    aes(y = after_stat(count))
    ) +
  scale_x_continuous(
    limits=c(-0.5,12.5),
    breaks=seq(0,12,by=2)
    ) +
  labs(
    x = "Perceptive Distance",
    y = "Number of Trials"
    ) +
  custom_theme()

dist
```

## Vocabulary richness

```{r vocab richness distribution, fig.height=5, fig.width=10}
# Word number distribution
WC <- ggplot(data, aes(x = nb_words)) +
  geom_histogram(
    binwidth = 1,
    fill=fillcolor,
    color="black",
    aes(y = after_stat(count))
    ) +
  labs(
    x = "Number of words",
    y = "Number of Trials"
    ) +
  custom_theme()

# Display plots
ggarrange(WC,
          nrow = 1,
          ncol = 1
)
```

## Semantic Distance

```{r lemma jaccard distribution, fig.height=5, fig.width=5}
# Semantic distance distribution
sem <- ggplot(data, aes(x = jaccard_dist)) +
  geom_histogram(
    binwidth = 0.1,
    fill=fillcolor,
    color="black",
    aes(y = after_stat(count))
    ) +
  scale_y_continuous(
  limits=c(0,600),
  breaks=seq(0,600,by=200)
  ) +
  labs(
    x = "Semantic Distance",
    y = "Number of Trials"
    ) +
  custom_theme()

sem
```

## Figure S1

```{r fig papier, fig.height=10, fig.width=20}
# Display plots
ggarrange(outcome_dist, pleasantness, emotional_strength, intensity,
          familiarity, dist, WC, sem,
          nrow = 2,
          ncol = 4
)
```

# Perceptual scores {.tabset .tabset-fade}

```{r perceptual variable selection}
# List of variables
variables <- c(
  "pleasantness",
  "emotional_strength",
  "intensity",
  "familiarity"
  )
```

## Variability

```{r plot rating variability, fig.height=20, fig.width=30}
# Reshape dataframe to long format
df_long <- df %>%
  gather(
    key = "Perceptual_variable",
    value = "Rating",
    pleasantness, intensity, familiarity, emotional_strength
    )

# Calculate averages
mean_ratings <- df_long %>%
  group_by(
    odor_name,
    Perceptual_variable
    ) %>%
  summarize(
    mean_rating = mean(Rating, na.rm = TRUE)
    )

# Reorder odor_name based on the mean pleasantness rating
mean_pleasantness <- mean_ratings %>%
  filter(
    Perceptual_variable == "pleasantness"
    ) %>%
  arrange(mean_rating)

df_long <- df_long %>%
  mutate(
    odor_name = factor(
      odor_name,
      levels = mean_pleasantness$odor_name
      )
    )

# Reorder Perceptual_variable
df_long <- df_long %>%
  mutate(Perceptual_variable = factor(
    Perceptual_variable,
    levels = c(
      "pleasantness",
      "emotional_strength",
      "intensity",
      "familiarity"
      )
    )
    )

mean_ratings <- mean_ratings %>%
  mutate(
    Perceptual_variable = factor(
      Perceptual_variable,
      levels = c(
        "pleasantness",
        "emotional_strength",
        "intensity",
        "familiarity"
        )
      )
    )

# Create graphics
ggplot(df_long, aes(x = odor_name, y = Rating)) +
  geom_line(
    aes(group = participant),
    alpha = 0.2,
    color = "grey",
    linewidth=0.5
    ) + 
  geom_point(
    alpha = 0.2,
    color = "grey",
    size=0.5
    ) +
  geom_line(
    data = mean_ratings,
    aes(x = odor_name, y = mean_rating, group = 1),
    color = "black",
    size = 1
    ) +
  facet_grid(
    ~ Perceptual_variable,
    scales = "free_x"
    ) +
  custom_theme() +
  labs(
    x= NULL,
    y = NULL
    ) +
  coord_flip()
```

## Correlations

```{r average by odor}
by_odor <- df %>%
  group_by(odor_name) %>%
  mutate(
    familiarity_u = ifelse(!is.na(unpleasant), familiarity, NA),
    familiarity_p = ifelse(!is.na(pleasant), familiarity, NA),
    intensity_u = ifelse(!is.na(unpleasant), intensity, NA),
    intensity_p = ifelse(!is.na(pleasant), intensity, NA)
  ) %>%
  summarize(
    pleasantness = mean(pleasantness, na.rm = TRUE),
    unpleasant = mean(unpleasant, na.rm = TRUE),
    pleasant = mean(pleasant, na.rm = TRUE),
    emotional_strength = mean(emotional_strength, na.rm = TRUE),
    intensity = mean(intensity, na.rm = TRUE),
    familiarity = mean(familiarity, na.rm = TRUE),
    familiarity_u = mean(familiarity_u, na.rm = TRUE),
    familiarity_p = mean(familiarity_p, na.rm = TRUE),
    intensity_u = mean(intensity_u, na.rm = TRUE),
    intensity_p = mean(intensity_p, na.rm = TRUE)
  )
```

```{r corr plot var selection}
# Define variable pairs
combinations <- list(
  list("intensity", "familiarity"),
  list("unpleasant", "familiarity_u"),
  list("pleasant", "familiarity_p"),
  list("intensity", "emotional_strength"),
  list("unpleasant", "intensity_u"),
  list("pleasant", "intensity_p")
)

# Define axis limits
axis_limits <- list(
  "pleasantness" = c(-5, 5),
  "unpleasant" = c(-5, 0),
  "pleasant" = c(0, 5),
  "emotional_strength" = c(0, 5),
  "intensity" = c(0, 10),
  "intensity_u" = c(0, 10),
  "intensity_p" = c(0, 10),
  "familiarity" = c(0, 10),
  "familiarity_u" = c(0, 10),
  "familiarity_p" = c(0, 10)
)
```

```{r corr plot, fig.height=10, fig.width=15}
# List to store raw p-values before correction
p_values <- numeric(length(combinations))

# List to store plots
gg_list <- list()

# Function to format axis labels
format_axis_labels <- function(var) {
  gsub("_u|_p", "", var)
}

# Function to capitalize words
format_to_title_case <- function(var) {
  tools::toTitleCase(var)
}

# Loop to compute correlations and store raw p-values
for (i in seq_along(combinations)) {
  combo <- combinations[[i]]
  x_var <- combo[[1]]
  y_var <- combo[[2]]

  # Format axis labels
  axis_labels <- list("emotional_strength" = "Emotional strength")
  formatted_x_var <- axis_labels[[x_var]] %||% format_to_title_case(format_axis_labels(x_var))
  formatted_y_var <- axis_labels[[y_var]] %||% format_to_title_case(format_axis_labels(y_var))

  # Compute correlation test
  cor_test <- cor.test(
    by_odor[[x_var]],
    by_odor[[y_var]],
    use = "complete.obs"
    )
  correlation <- cor_test$estimate
  p_values[i] <- cor_test$p.value

  # Generate plot name
  plot_name <- paste(
    x_var,
    y_var,
    sep = "_vs_"
    )

  # Create the plot (p-value not yet corrected)
  gg_list[[plot_name]] <- ggplot(
    by_odor,
    aes_string(x = x_var, y = y_var)
    ) +
    geom_point(
      size = 2,
      color = "black"
      ) +
    geom_smooth(
      method = "lm",
      se = FALSE,
      color = "gray"
      ) +
    labs(
      x = formatted_x_var,
      y = formatted_y_var
      ) +
    scale_x_continuous(
      limits = axis_limits[[x_var]]
      ) +
    scale_y_continuous(
      limits = axis_limits[[y_var]]
      ) +
    custom_theme()
}

# Apply p-value correction
p_values_corr <- p.adjust(p_values, method = "holm")

# Update plots with corrected p-values
for (i in seq_along(combinations)) {
  combo <- combinations[[i]]
  x_var <- combo[[1]]
  y_var <- combo[[2]]
  plot_name <- paste(
    x_var,
    y_var,
    sep = "_vs_"
    )

  # Generate updated label with corrected p-value
  correlation <- cor.test(
    by_odor[[x_var]],
    by_odor[[y_var]],
    use = "complete.obs")$estimate
  label <- paste(
    "r =",
    format(
      round(correlation, 4),
      nsmall = 4
      ),
    "\np =",
    ifelse(
      p_values_corr[i] < 1e-4,
      formatC(
        p_values_corr[i],
        format = "e",
        digits = 2
        ),
      format(
        round(p_values_corr[i], 4),
        nsmall = 4
        )
      )
  )

  # Add annotation with corrected p-value
  gg_list[[plot_name]] <- gg_list[[plot_name]] +
    annotate(
      "text",
      x = Inf,
      y = -Inf,
      label = label,
      hjust = 1,
      vjust = -0.5,
      size = 6,
      color = "black"
      )
}

# Arrange plots
ggarrange(
  gg_list[["intensity_vs_familiarity"]],
  gg_list[["unpleasant_vs_familiarity_u"]],
  gg_list[["pleasant_vs_familiarity_p"]],
  gg_list[["intensity_vs_emotional_strength"]],
  gg_list[["unpleasant_vs_intensity_u"]],
  gg_list[["pleasant_vs_intensity_p"]],
  ncol = 3, nrow = 2
)
```

# Gender x Memory scores {.tabset .tabset-fade}

## Gender 

```{r outcome x gender}
# Define model formulas
model_formulas <- list(
  outcome ~ gender_encoded + (1 | participant) + (1 | study / odor_name),
  outcome ~ gender_encoded + (1 | participant) + (1 | study) + (1 | odor_name),
  outcome ~ gender_encoded + (1 | participant)
)

# Adjust models
models <- lapply(model_formulas, function(formula) {
  glmmTMB(
    formula,
    family = binomial,
    data = data
    )
})

# Naming models for easy identification
names(models) <- paste0("model", 1:length(models))

# Model comparison by AIC
model_selection <- model.sel(models)
best_model <- get.models(model_selection, 1)[[1]]

# Results
print(model_selection)
summary(best_model)
```

## Chance level

```{r memory prop x chance level}
# Stats by subj
by_subj %>%
  group_by(gender) %>%
  t_test(
    outcome ~ 1,
    mu = chance,
    alternative = "greater"
    ) %>%
  adjust_pvalue(method = "holm") %>%
  add_significance() %>%
  kbl(caption = "By subj - One-tailed t-test, Holm correction") %>%
  kable_classic(full_width = F, html_font = "Cambria")

# Stats by trial
data %>%
  group_by(gender) %>%
  t_test(
    outcome ~ 1,
    mu = chance,
    alternative = "greater"
    ) %>%
  adjust_pvalue(method = "holm") %>%
  add_significance() %>%
  kbl(caption = "By trial - One-tailed t-test, Holm correction") %>%
  kable_classic(full_width = F, html_font = "Cambria")
```

# Gender x features

```{r random structures}
# Random structures to test
random_structures <- list(
  "(1 | participant) + (1 | study / odor_name)",
  "(1 | participant) + (1 | study) + (1 | odor_name)",
  "(1 | participant)"
)
```

```{r gender x features}
# List of variables and their respective models
variables_lmer <- c(
  "pleasantness",
  "emotional_strength",
  "intensity",
  "familiarity",
  "percept_dist",
  "jaccard_dist"
  )

variables_glmmTMB_nbinom2 <- c("nb_words")

# List to store the best models
best_models <- list()

# Function to fit models based on variable type
fit_model <- function(variable, formula, data) {
  if (variable %in% variables_lmer) {
    return(tryCatch(
      lmer(
        formula,
        data = data,
        control = lmerControl(
          check.conv.grad = .makeCC(
            "ignore",
            tol = 0.002
            )
          )
        ),
      error = function(e) NULL
    )
    )
    
  } else if (variable %in% variables_glmmTMB_nbinom2) {
    return(tryCatch(
      glmmTMB(
        formula,
        data = data,
        family = nbinom2
        ),
      error = function(e) NULL
    ))
  }
  return(NULL)
}

# Loop for each variable
data <- data
for (variable in c(
  variables_lmer,
  variables_glmmTMB_nbinom2
  )
  ) {
  models <- list()
  
  # Adjust models for each random structure
  for (random_effects in random_structures) {
    if (variable == "pleasantness") {
      formula <- as.formula(
        paste0(
          variable,
          " + I(", variable, "^2) ~ gender_encoded + ",
          random_effects
          )
        )
    } else {
      formula <- as.formula(
        paste0(
          variable,
          " ~ gender_encoded + ",
          random_effects
          )
        )
    }
    
    model <- fit_model(
      variable,
      formula,
      data
      )
    
    if (!is.null(model)) {
      if (
        (variable %in% variables_lmer && !isSingular(model)
         ) || variable %in% c(
           variables_glmmTMB_nbinom2
           )
        ) {
        warnings <- tryCatch(
          summary(model)$optinfo$conv$lme4$messages,
          error = function(e) NULL
          )
        if (is.null(warnings) || !grepl(
          "boundary|Model failed to converge",
          paste(warnings),
          ignore.case = TRUE)
          ) {
          models[[random_effects]] <- model
        }
      }
    }
  }
  
  # Compare valid models by AIC
  if (length(models) > 0) {
    names(models) <- paste0(
      "model",
      1:length(models)
      )
    model_selection <- model.sel(models)
    best_models[[variable]] <- get.models(
      model_selection,
      1
      )[[1]]
  } else {
    best_models[[variable]] <- NULL
  }
}

# Display summaries of the best models
for (variable in c(
  variables_lmer,
  variables_glmmTMB_nbinom2)
  ) {
  cat("\nSummary for", variable, ":\n")
  if (!is.null(best_models[[variable]])) {
    print(summary(best_models[[variable]]))
  } else {
    cat("No valid model found for this variable.\n")
  }
  cat("\n\n--------------------------------------------------------------\n\n")
}

```

# Memory score x features

```{r include gender condition}
# List of variables
variables <- c(
  "gender_encoded",
  "pleasantness",
  "emotional_strength",
  "intensity",
  "familiarity",
  "percept_dist",
  "nb_words",
  "jaccard_dist"
  )

include_gender <- if (
  condition == "odor_recognition"
  ) {
  "+ gender_encoded"
} else if (
  condition == "associative_memory" | "gender_encoded" %in% variables
  ) {
  ""
} else {
  ""
}
```

```{r outcome x features, fig.height=10, fig.width=20}
# Function to format p-value
format_p_value <- function(p) {
  if (p < 0.001) {
    return(format(p, scientific = TRUE, digits = 2))
  } else {
    return(sprintf("%.3f", p))
  }
}

# List for storing model results and plots
model_list <- list()
gg_list <- list()

# Loop to fit models for each variable
for (variable in variables) {
  # Initialize a list to store models for the current variable
  models_for_variable <- list()

  # For 'pleasantness', fit only the quadratic model
  if (variable == "pleasantness") {
    formula_quadratic <- as.formula(
      paste0(
        "outcome ~ ",
        variable,
        " + I(", variable, "^2) ",
        include_gender
        )
      )
    for (random_structure in random_structures) {
      formula_with_random <- paste(
        deparse(formula_quadratic),
        random_structure,
        sep = " + "
        )
      model_quadratic <- glmmTMB(
        as.formula(formula_with_random),
        family = binomial,
        data = data
        )
      models_for_variable[[paste0(
        variable,
        "_quadratic_",
        random_structure
        )
        ]
        ] <- model_quadratic
    }
  }

  # For 'familiarity' fit both the linear and quadratic models
  else if (variable == "familiarity") {
    # Linear model
    formula_linear <- as.formula(
      paste0(
        "outcome ~ ",
        variable,
        include_gender
        )
      )
    for (random_structure in random_structures) {
      formula_with_random <- paste(
        deparse(formula_linear),
        random_structure,
        sep = " + "
        )
      model_linear <- glmmTMB(
        as.formula(formula_with_random),
        family = binomial,
        data = data
        )
      models_for_variable[[paste0(
        variable,
        "_linear_",
        random_structure
        )
        ]
        ] <- model_linear
    }

    # Quadratic model
    formula_quadratic <- as.formula(
      paste0(
        "outcome ~ ",
        variable,
        " + I(", variable, "^2) ",
        include_gender
        )
      )
    for (random_structure in random_structures) {
      formula_with_random <- paste(
        deparse(formula_quadratic),
        random_structure,
        sep = " + "
        )
      model_quadratic <- glmmTMB(
        as.formula(formula_with_random),
        family = binomial,
        data = data
        )
      models_for_variable[[paste0(
        variable,
        "_quadratic_",
        random_structure
        )
        ]
        ] <- model_quadratic
    }
  }

  # For other variables, fit only the linear model
  else {
    formula_linear <- as.formula(
      paste0(
        "outcome ~ ",
        variable,
        include_gender
        )
      )
    for (random_structure in random_structures) {
      formula_with_random <- paste(
        deparse(formula_linear),
        random_structure,
        sep = " + "
        )
      model_linear <- glmmTMB(
        as.formula(formula_with_random),
        family = binomial,
        data = data
        )
      models_for_variable[[paste0(
        variable,
        "_linear_",
        random_structure
        )
        ]
        ] <- model_linear
    }
  }

  # Get the best model by AIC
  best_model <- model.sel(
    models_for_variable
    )[which.min(
      model.sel(
        models_for_variable
        )$AIC
      ),
      ]
  best_model_name <- rownames(best_model)
  best_model_fit <- models_for_variable[[best_model_name]]

  # Extract z and p-values for the variable
  summary_best_model <- summary(best_model_fit)

  if (variable %in% c("pleasantness", "familiarity") && grepl("quadratic", best_model_name)) {
    quadratic_term <- paste0("I(", variable, "^2)")
    variable_coef <- summary_best_model$coefficients$cond[quadratic_term, ]
  } else {
    variable_coef <- summary_best_model$coefficients$cond[variable, ]
  }
  
  z_value <- round(variable_coef["z value"], 3)
  p_value <- as.numeric(format_p_value(variable_coef["Pr(>|z|)"]))

  # Store the best model summary for the current variable
  model_list[[variable]] <- summary_best_model

  # Rename axis labels
  variable_label <- switch(
    variable,
    "gender_encoded" = "Gender",
    "nb_words" = "Number of words",
    "percept_dist" = "Perceptive distance",
    "jaccard_dist" = "Semantic distance",
    "emotional_strength" = "Emotional strength",
    paste0(
      toupper(substr(variable, 1, 1)),
      substr(variable, 2, nchar(variable)))
    )

  # Create the plot for the best model
  if (variable == "gender_encoded") {
    data$prediction_linear <- predict(
      best_model_fit,
      type = "response"
      )
    gg_hit <- ggplot(data,
                     aes(
                       x = factor(!!sym(variable)),
                       y = prediction_linear * 100)
                     ) +
      geom_boxplot(
        fill = fillcolor,
        color = "black",
        size = 1
        ) +
      stat_summary(
        fun.data = mean_cl_normal,
        geom = "errorbar",
        width = 0.2,
        color = "red",
        linewidth = 1
        ) +
      stat_summary(
        fun = mean,
        geom = "point",
        shape = 23,
        size = 3,
        fill = "red"
        ) +
      scale_x_discrete(
        labels = if (variable == "gender_encoded") c("W", "M")
        ) +
      scale_y_continuous(
        limits = c(0, 106)
        ) +
      labs(
        x = variable_label,
        y = "Predicted Probability \n to Succeed (%)"
      ) +
      custom_theme() +
      annotate(
        "text",
        x = Inf,
        y = -Inf,
        label = paste0(
          "z = ", round(z_value, 3),
          "\np = ", round(p_value, 3)
          ),
        hjust = 1.1,
        vjust = -0.5,
        size = 7
        )
    
    if (variable == "gender_encoded") {
      gg_hit <- gg_hit + geom_hline(
        yintercept = chance * 100,
        linetype = "dashed",
        color = "red",
        linewidth = 1
        )
    }
    
  } else if (grepl("linear", best_model_name)) {
    data$prediction_linear <- predict(
      best_model_fit,
      type = "response"
      )
    gg_hit <- ggplot(
      data,
      aes_string(
        x = variable,
        y = "prediction_linear * 100")
      ) +
      geom_point(size = 1) +
      labs(
        x = variable_label,
        y = "Predicted Probability \n to Succeed (%)"
      ) +
      custom_theme() +
      scale_x_continuous(
        limits = if (variable == "percept_dist") c(0, 12) else NULL,
        breaks = if (variable == "percept_dist") seq(0, 12, by = 2) else if (variable == "nb_words") seq(0, 12, by = 1) else waiver(),
        labels = if (variable == "nb_words") function(x) ifelse(
          x %% 2 == 0,
          as.character(x),
          "") else waiver(),
        expand = expansion(mult = c(0.04, 0.08))
                         ) +
      scale_y_continuous(limits = c(0, 106)) +
      geom_smooth(method = "glm",
                  formula = y ~ x,
                  color = fillcolor,
                  se = FALSE,
                  linewidth = 1) +
      annotate(
        "text",
        x = Inf,
        y = -Inf,
        label = paste0(
          "z = ", round(z_value, 3),
          "\np = ", p_value),
        hjust = 1.1,
        vjust = -0.5,
        size = 7
        )
    
  } else if (grepl("quadratic", best_model_name)) {
    data$prediction_quadratic <- predict(
      best_model_fit,
      type = "response"
      )
    gg_hit <- ggplot(
      data,
      aes_string(
        x = variable,
        y = "prediction_quadratic * 100")
      ) +
      geom_point(size = 1) +
      labs(
        x = variable_label,
        y = "Predicted Probability \n to Succeed (%)"
      ) +
      custom_theme() +
      scale_x_continuous(expand = expansion(mult = c(0.04, 0.08))) +
      scale_y_continuous(
        limits = c(0, 106)
        ) +
      geom_smooth(
        method = "glm",
        formula = y ~ poly(x, 2),
        color = fillcolor,
        se = FALSE,
        linewidth = 1
        ) +
      annotate(
        "text",
        x = Inf,
        y = -Inf,
        label = paste0(
          "z = ", round(z_value, 3),
          "\np = ", p_value),
        hjust = 1.1,
        vjust = -0.5,
        size = 7
        )
  }

  # Store the plot in the gg_list
  gg_list[[variable]] <- gg_hit
}

# Display only the best model for each variable
for (variable in variables) {
  cat("\nSummary for", variable, ":\n")
  if (!is.null(model_list[[variable]])) {
    print(model_list[[variable]])
  } else {
    cat("No valid model found for this variable.\n")
  }
  cat("\n\n--------------------------------------------------------------\n\n")
}

# Combine and display all plots
ggarrange(
  plotlist = gg_list,
  nrow = 2, ncol = 4
)
```